### 2018-08-10 NYC_Twitter_SDM_models.Rmd
Upload files to repository
New park access analysis with data and code.

---
title: "NYC Spatial Models - Comparison of SDM Models With and Without Twitter Data"
author: "Nico Boyd"
date: "June 26, 2017"
output: html_document
---

For both obesity and physical activity, the first SDM model uses a size beta of .00001 and a distance beta of -5. The second SDM model uses the same size and distance betas with the addition of a tweet beta of .001. The third SDM model again uses the same size and distance betas but uses a negative tweet beta of -.001.

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(geojsonio)
library(tidyverse)
library(maptools)
library(rgdal)
library(rgeos)
library(spdep)
library(texreg)
library(leaflet)
```

```{r read_tracts, echo=FALSE}
tracts <- geojson_read("nyc_tracts.geojson", what = "sp")
#plot(tracts)
```

```{r read_parks, echo=FALSE}
parks <- geojson_read("nyc_parks.geojson", what = "sp")
#plot(parks)
```

```{r show_data, echo=FALSE}
tracts@data <- tbl_df(tracts@data) %>%
  select(GEOID, OBESITY, Park_Percent, Phys_Act, MENTAL, Income1, Income2, Income3, Income4, Income5, Income6, Income7, Income8, Income9, Income10, Pop_Density, FulltimeWork, CollegeDeg, Pct0to17, Pct18to29, Pct30to64, Pct65plus, Single_Percent, PctWhite, PctBlack, PctNative, PctAsian, PctPacific, PctOther, PctHispanic) %>%
    mutate_each(funs(ifelse(is.na(.), 0, .)), Park_Percent)

parks@data <- tbl_df(parks@data)

tracts@data
#parks@data
```

Calculate the logsum given a tracts and parks dataset. `tracts$park_ls1` uses only size and distance parameters. `tracts$park_ls2` uses size, distance and tweet parameters. `tracts$park_ls3` uses a size and distance parameter with a negative tweet parameter.
```{r accessibility}
source('size00001dist5.R', echo=TRUE)
tracts$park_ls1 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -5))

source('size00001dist5tweet001.R', echo=TRUE)
tracts$park_ls2 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -5, .001))

source('size00001dist5tweetneg001.R', echo=TRUE)
tracts$park_ls3 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -5, -.001))

source('size00001dist15.R', echo=TRUE)
tracts$park_ls4 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -15))

source('size00001dist15tweet001.R', echo=TRUE)
tracts$park_ls5 <- calculate_park_logsums(tracts, parks, betas = c(.00001, -15, .001))
```

Map the newly created Park Access variables. Here is a map of accessibility using only the size and distance parameters that are most likely when using obesity as the dependent variable.
```{r accessibility_leaflet_1, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls1)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls1, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls1))
```

And here is a map of accessibility using the size, distance and tweet parameters (using obesity as the DV). The tweet parameter here is positive.
```{r accessibility_leaflet_2, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls2)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls2, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls2))
```

Here is a map of accessibility using the betas that produce the most likely model when using physical activity as the dependent variable. Only size and distance parameters are used.
```{r accessibility_leaflet_3, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls4)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls4, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls4))
```

And here is a map of accessibility using size, distance and tweet parameters (with the combination of betas that is most likely for physical activity as the DV).
```{r accessibility_leaflet_4, echo=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls5)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls5, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls5))
```

```{r accessibility_leaflet_X, echo=FALSE, eval=FALSE}
# Create a continuous palette function
pal <- colorQuantile(
  palette = "Greens",
  domain = c(0, tracts$park_ls3)
)

pop_content <- function(shp){
  paste0(
    "Tract: ", shp$NAME, "<br/>",
    "% Park: ", round(shp$Park_Percent, 2), "<br/>",
    "Park Access: ", round(shp$park_ls3, 2), "<br/>"
  )
}

leaflet(tracts %>% spTransform(., CRS("+init=epsg:4326"))) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(weight = 1, popup = ~pop_content(tracts),
              color = ~pal(park_ls3))
```

```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
base_formula <- formula(~log(Pop_Density) + FulltimeWork + 
    CollegeDeg + Single_Percent + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)
base_formula_age <- formula(~log(Pop_Density) + FulltimeWork + 
    CollegeDeg + Single_Percent + Pct18to29 + Pct30to64 + Pct65plus + PctWhite + PctBlack + PctAsian + PctOther + PctHispanic)

parks_formula_1 <- update(base_formula, . ~ . + log(park_ls1))
parks_formula_1a <- update(base_formula_age, . ~ . + log(park_ls1))
parks_formula_2 <- update(base_formula, . ~ . + log(park_ls2))
parks_formula_2a <- update(base_formula_age, . ~ . + log(park_ls2))
parks_formula_3 <- update(base_formula, . ~ . + log(park_ls4))
parks_formula_3a <- update(base_formula_age, . ~ . + log(park_ls4))
parks_formula_4 <- update(base_formula, . ~ . + log(park_ls5))
parks_formula_4a <- update(base_formula_age, . ~ . + log(park_ls5))

#parks_formula_3 <- update(base_formula, . ~ . + park_ls3)
```

Create a spatial weights matrix:

```{r W, echo=FALSE}
neighbors <- poly2nb(tracts, queen = FALSE)
W <- nb2listw(neighbours = neighbors, zero.policy = TRUE)
print(W, zero.policy = TRUE)
```

Estimate SDM models for the `OBESITY` variable. Model 1 does not include park access. Models 2 & 3 each use a different park access variable with a different combination of parameters:

```{r spmodels obesity, echo=FALSE}
obese_sdm <- lagsarlm(update(base_formula, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdm2 <- lagsarlm(update(parks_formula_1, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdm3 <- lagsarlm(update(parks_formula_2, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdem <- errorsarlm(update(base_formula, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
obese_sdem2 <- errorsarlm(update(parks_formula_1, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
obese_sdem3 <- errorsarlm(update(parks_formula_2, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
#obese_sdm_4 <- lagsarlm(update(parks_formula_3, OBESITY ~ .), data = tracts@data, 
                      #listw = W, zero.policy = TRUE, type = "mixed")
```

```{r spmodels obesity with age, echo=FALSE}
obese_sdm_age <- lagsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdm2_age <- lagsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdm3_age <- lagsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, type = "mixed")
obese_sdem_age <- errorsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
obese_sdem2_age <- errorsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
obese_sdem3_age <- errorsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)

```

Estimate SDM models for the `Phys_Act` variable. Model 1 does not include park access. Models 2-4 each use a different park access variable with a different combination of parameters:

```{r spmodels physact, echo=FALSE}
physact_sdm <- lagsarlm(update(base_formula, Phys_Act ~ .), data = tracts@data, 
                        listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm2 <- lagsarlm(update(parks_formula_1, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm3 <- lagsarlm(update(parks_formula_2, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdem <- errorsarlm(update(base_formula, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem2 <- errorsarlm(update(parks_formula_1, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem3 <- errorsarlm(update(parks_formula_2, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
#physact_sdm4 <- lagsarlm(update(parks_formula_3, Phys_Act ~ .), data = tracts@data,
                         #listw = W, zero.policy = TRUE, type = "mixed")
```

```{r spmodels physact with age, echo=FALSE}
physact_sdm_age <- lagsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
                        listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm2_age <- lagsarlm(update(parks_formula_1a, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm3_age <- lagsarlm(update(parks_formula_2a, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdem_age <- errorsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem2_age <- errorsarlm(update(parks_formula_1a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem3_age <- errorsarlm(update(parks_formula_2a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)

```

```{r spmodels physact new betas, echo=FALSE}
physact_sdm_new <- lagsarlm(update(base_formula, Phys_Act ~ .), data = tracts@data, 
                        listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm2_new <- lagsarlm(update(parks_formula_3, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm3_new <- lagsarlm(update(parks_formula_4, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdem_new <- errorsarlm(update(base_formula, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem2_new <- errorsarlm(update(parks_formula_3, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem3_new <- errorsarlm(update(parks_formula_4, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
```

```{r spmodels physact age new betas, echo=FALSE}
physact_sdm_new_age <- lagsarlm(update(base_formula_age, Phys_Act ~ .), data = tracts@data, 
                        listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm2_new_age <- lagsarlm(update(parks_formula_3a, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdm3_new_age <- lagsarlm(update(parks_formula_4a, Phys_Act ~ .), data = tracts@data,
                         listw = W, zero.policy = TRUE, type = "mixed")
physact_sdem_new_age <- errorsarlm(update(base_formula_age, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem2_new_age <- errorsarlm(update(parks_formula_3a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
physact_sdem3_new_age <- errorsarlm(update(parks_formula_4a, OBESITY ~ .), data = tracts@data, 
                      listw = W, zero.policy = TRUE, etype = "emixed",tol.solve = 1e-16)
```
Compare the models with `htmlreg()` in
the html output. First, the obesity models. The first table does not include the age variables.

```{r obesitysummary, results="asis", echo=FALSE}
obesity_models <- list(obese_sdm, obese_sdm2, obese_sdm3, obese_sdem, obese_sdem2, obese_sdem3)
names(obesity_models) <-  paste(c("SDM No Parks", "SDM Size & Distance", "SDM Size, Distance, Tweets", "SDEM 1", "SDEM 2", "SDEM 3"))
#screenreg(obesity_models, digits = 4)
htmlreg(obesity_models, caption = "NYC: SDM & SDEM Models of Obesity Prevalence")
```

```{r obesitysummaryage, results="asis", echo=FALSE}
obesity_models_age <- list(obese_sdm_age, obese_sdm2_age, obese_sdm3_age, obese_sdem_age, obese_sdem2_age, obese_sdem3_age)
names(obesity_models_age) <-  paste(c("SDM No Parks", "SDM Size & Distance", "SDM Size, Distance, Tweets", "SDEM 1", "SDEM 2", "SDEM 3"))
#screenreg(obesity_models, digits = 4)
htmlreg(obesity_models_age, caption = "NYC: SDM & SDEM Models of Obesity Prevalence")
```

Estimate the impacts for `obese_sdm`.
```{r obese_sdm_impacts}

listw <- as(W, "CsparseMatrix")
trMC <- trW(listw, type="MC")
set.seed(1)
summary(impacts(obese_sdm, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `obese_sdm_age`.
```{r obese_sdm_age_impacts}

listw <- as(W, "CsparseMatrix")
trMC <- trW(listw, type="MC")
set.seed(1)
summary(impacts(obese_sdm_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `obese_sdm2`.
```{r obese_sdm2_impacts}

summary(impacts(obese_sdm2, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `obese_sdm2_age`.
```{r obese_sdm2_age_impacts}

summary(impacts(obese_sdm2_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `obese_sdm3`.
```{r obese_sdm3_impacts}

summary(impacts(obese_sdm3, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `obese_sdm3_age`.
```{r obese_sdm3_age_impacts}

summary(impacts(obese_sdm3_age, tr=trMC, R=1000), zstats=TRUE)

```

Next, the physical activity models. Again, the first table does not contain the age variables. The third and fourth tables use new beta parameters.

```{r physactsummary, results="asis", echo=FALSE}
physact_models <- list(physact_sdm, physact_sdm2, physact_sdm3, physact_sdem, physact_sdem2, physact_sdem3)
names(physact_models) <-  paste(c("SDM No Parks", "SDM Size & Distance", "SDM Size, Distance, Tweets", "SDEM 1", "SDEM 2", "SDEM 3"))
#screenreg(physact_models)
htmlreg(physact_models, caption = "NYC: SDM Models of Physical Activity Participation")
```

```{r physactsummaryage, results="asis", echo=FALSE}
physact_models_age <- list(physact_sdm_age, physact_sdm2_age, physact_sdm3_age, physact_sdem_age, physact_sdem2_age, physact_sdem3_age)
names(physact_models_age) <-  paste(c("SDM No Parks", "SDM Size & Distance", "SDM Size, Distance, Tweets", "SDEM 1", "SDEM 2", "SDEM 3"))
#screenreg(physact_models)
htmlreg(physact_models_age, caption = "NYC: SDM Models of Physical Activity Participation")
```

The following two tables use a size beta of 0.00001 and a distance beta of -15.
```{r physact new betas, results="asis", echo=FALSE}
physact_models_new <- list(physact_sdm_new, physact_sdm2_new, physact_sdm3_new, physact_sdem_new, physact_sdem2_new, physact_sdem3_new)
names(physact_models_new) <-  paste(c("SDM No Parks", "SDM Size & Distance", "SDM Size, Distance, Tweets", "SDEM 1", "SDEM 2", "SDEM 3"))
#screenreg(physact_models)
htmlreg(physact_models_new, caption = "NYC: SDM Models of Physical Activity Participation")

```

```{r physact new betas with age, results="asis", echo=FALSE}
physact_models_new_age <- list(physact_sdm_new_age, physact_sdm2_new_age, physact_sdm3_new_age, physact_sdem_new_age, physact_sdem2_new_age, physact_sdem3_new_age)
names(physact_models_new_age) <-  paste(c("SDM No Parks", "SDM Size & Distance", "SDM Size, Distance, Tweets", "SDEM 1", "SDEM 2", "SDEM 3"))
#screenreg(physact_models)
htmlreg(physact_models_new_age, caption = "NYC: SDM Models of Physical Activity Participation")

```

Estimate the impacts for `physact_sdm`.
```{r physact_sdm_impacts}

summary(impacts(physact_sdm, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm_age`.
```{r physact_sdm_age_impacts}

summary(impacts(physact_sdm_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm2`.
```{r physact_sdm2_impacts}

summary(impacts(physact_sdm2, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm2_age`.
```{r physact_sdm2_age_impacts}

summary(impacts(physact_sdm2_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm3`.
```{r physact_sdm3_impacts}

summary(impacts(physact_sdm3, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm3_age`.
```{r physact_sdm3_age_impacts}

summary(impacts(physact_sdm3_age, tr=trMC, R=1000), zstats=TRUE)

```

Estimate the impacts for `physact_sdm3_new_age`.
```{r physact_sdm3_new_age_impacts}

summary(impacts(physact_sdm3_new_age, tr = trMC, R = 1000), zstats=TRUE)

```