### 2018-12-07 model_estimation.Rmd
Fix bug where park/tract distance matrix was built inverted

---
title: "Model Estimation"
output:
  html_document:
    df_print: paged
editor_options: 
  chunk_output_type: console
---
```{r setup}
library(geojsonio)
library(tidyverse)
library(maptools)
library(rgdal)
library(rgeos)
library(spdep)
library(texreg)
library(leaflet)
library(sf)
library(VGAM)
```

# Data cleaning

Nico cleaned the tracts dataset into a geoJSON file stored in the repository. We
should read it in twice because the sf library is better for visualization but
spdep still requires sp objects for now.

```{r read_tracts}
tracts <- geojson_read("data/nyc_tracts.geojson", what = "sp")
tracts_sf <- st_read("data/nyc_tracts.geojson", quiet = TRUE) %>%
  st_transform(4326)
```

## Complete Cases
We noticed that our models had different numbers of observations / degrees of 
freedom. Obviously we need to have comparable econometric models. The first thing
we want to look for is missingness in the tracts variables, and which tracts
are missing those variables. The map below shows the tracts that are missing
this information.

```{r tracts_with_missing}
# identify complete cases
complete_tracts_index <- tracts_sf %>%
  st_set_geometry(NULL) %>%
  tbl_df() %>%
  complete.cases()

# make a map of 
leaflet( tracts_sf[!complete_tracts_index, ]) %>%
  addProviderTiles(providers$OpenStreetMap.Mapnik) %>% 
  addPolygons(label = ~ as.character(GEOID))
```

These tracts seem to fall into two basic categories:

  - Tracts outside of New York City that were clipped into the shape layer 
  and remain as edge remnants. It is generally better to select by FIPS code rather
  than geometric operations, but we can just discard these at this point and
  they will not affect anything.

  - Tracts that are completely occupied by parks, airports, or other pieces
  of non-residential infrastructure.

The exceptions to this appear to be 3606100940 and 36061010200 in Midtown between
5th Avenue and Park Avenue. Greg will assume for this analysis that these are 
entirely commercial tracts and therefore have no residents, but Nico needs to 
verify these.

```{r complete_cases}
tracts <- tracts[complete_tracts_index, ]
```

The parks data seems relatively complete, with the exception of a few variables
that are missing everywhere. But we'll ignore this for now.

```{r read_parks, echo=FALSE}
parks <- sf::st_read("data/nyc_parks.geojson")
```

## Tweets Transformation

The Tweets count variable has a high degree of zeros (virtually all parks have
no tweets originating there). We should identify if there is a transformation
that will make the distribution more normal (log is typical), but that will allow
for zero variables as $log(0)$ is undefined. Using a Yeo-Johnson transformation
with a lambda value of zero transforms the tweet count into a distribution
asymptotically equivalent to log, and with zero values equal to zero.

```{r tweettransformation}
parks %>%
  tbl_df() %>%
  dplyr::select(raw = TWEET_COUNT) %>%
  filter(raw < 100) %>%
  mutate(
    `0.5` = yeo.johnson(raw, 0.5),
    `0` = yeo.johnson(raw, 0),
    log = log(raw)
  ) %>%
  gather(lambda, x, -raw) %>%
  ggplot(aes(x = raw, y = x, color = lambda)) + 
  geom_line()

parks <- parks %>%
  mutate(
    log_size = log(Park_Acres),
    log_tweets = yeo.johnson(TWEET_COUNT, 0)
  )

```

## Other variables

```{r show_data, echo=FALSE}
tracts@data <- tbl_df(tracts@data) %>%
  dplyr::select(
    GEOID, OBESITY, Park_Percent, Phys_Act, MENTAL, 
    Income1, Income2, Income3, Income4, Income5, Income6, 
    Income7, Income8, Income9, Income10, Pop_Density, 
    FulltimeWork, CollegeDeg, Pct0to17, Pct18to29, Pct30to64, Pct65plus, 
    Single_Percent, PctWhite, PctBlack, PctNative, PctAsian, PctPacific,
    PctOther, PctHispanic
  )  %>%
  mutate(
    log_obesity = log(OBESITY),
    log_physact = log(100 - Phys_Act)
  )

```

# Spatial Model Selection

Deciding which spatial model to select is an empirical question as outlined by
both LeSage and Pace (2010?) and Macfarlane et al. (2015) We estimate a model
without park accessibility to determine which spatial structure is correct
for our analysis.

We assert a first-order queen adjacency neighbors matrix, which is
row-standardized such that the total weight of each row in the matrix sums to 1.
There are three tracts in the data that have no adjacent neighbors, meaning that 
these tracts will have no spatial spillover effects.

```{r W, echo=FALSE}
tracts.dnn <- dnearneigh(gCentroid(tracts, byid = TRUE), 0, 1.8 * 5280)
dists <- nbdists(tracts.dnn, gCentroid(tracts, byid = TRUE))
dists.inv <- lapply(dists, function(x) 1 / x)
W <- nb2listw(neighbours = tracts.dnn, glist = dists.inv, zero.policy = TRUE, style = "W")
print(W, zero.policy = TRUE)
trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
```

The base model regresses the obesity rate in the tract against covariates for
work status and educational attainment, minority and young/elderly population, 
and marital status.

```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
base_formula <- formula( 
  ~ log(Pop_Density) + 
    FulltimeWork + CollegeDeg + Single_Percent + 
    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity  
    PctBlack + PctAsian + PctOther + PctHispanic)
```

We estimate four models with this regression specification:

  - Ordinary least squares
  - Spatial autoregressive
  - Spatial error
  - Spatial Durbin

```{r spatial selection}
obese_base_lm <- lm(update(base_formula, log_obesity ~ .), data = tracts@data)
obese_base_sar <- lagsarlm(update(base_formula, log_obesity ~ .), 
                           data = tracts@data, listw = W, zero.policy = TRUE)
obese_base_sem <- errorsarlm(update(base_formula, log_obesity ~ .), 
                             data = tracts@data, listw = W, zero.policy = TRUE)
obese_base_sdm <- lagsarlm(update(base_formula, log_obesity ~ .), 
                           data = tracts@data, listw = W, zero.policy = TRUE, 
                           type = "mixed")
```

A likelihood ratio test reveals that the SEM is not preferred, so we use the SDM
only going forward.

```{r lrtest}
lmtest::lrtest(obese_base_sem, obese_base_sdm)
```

The residuals appear homoskedastic with few outliers.

```{r residuals}
tracts_sf <- tracts_sf[complete_tracts_index, ] %>%
  mutate(
    log_obesity = tracts$log_obesity,  
    obese_base_lm_residuals = obese_base_lm$residuals,
    obese_base_sdm_fitted = obese_base_sdm$fitted.values,  
    obese_base_sdm_residuals = obese_base_sdm$residual
  )

ggplot(tracts_sf, aes(x = log_obesity, y = obese_base_sdm_fitted)) + 
  geom_point()
```

The residuals also appear to be distributed wtihout serious spatial correlation
```{r residuals_map}
pal <- colorQuantile("BrBG", tracts_sf$obese_base_sdm_residuals, n = 9)
leaflet(tracts_sf) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(color = ~pal(obese_base_sdm_residuals), stroke = NA, opacity = 0.8)
```

# Logsum Calculation

Consider that an individual is choosing a park for a recreation activity.
According to basic choice theory (Mcfadden 1974), the probability of choosing
park $p$ from the set of all regional parks $J$ is:

$$ P(p | V_p) = \frac{\exp(V_p)}{\sum_{j \in J}\exp(V_j)}$$
where parks are differentiated from each other by their relative measurable
utilitie $V$. In principle, $V$ may include any measurable attributes of either
the choice maker or the park. In this study we use a linear formulation of

$$V_{ij} = size_j\beta_s + distance_{ij}\beta_d$$
incorporating the size of the park in acres and the distance of the park from
the census tract $i$ in miles. The coefficients $\beta$ are typically estimated from 
surveys, though in the absence of a survey we apply a maximum likelihood
technique described below.

A key theoretical understanding of random utility choice models is that the
consumer surplus of the choice set can be otained as the log-sum of the
denominator of the choice probability equation. In plainer terms, the *total value*
of an individual's park accessibility is defined to be:

$$CS_i = \ln\left({\sum_{j \in J}\exp(V_{ij})}\right)$$

There are several advantages to a log-sum defined metric relative to
buffer-based accessibility metrics more commonly found in the literature. First,
all individuals are defined as having some access to all parks, rather than an
arbitrary limit of 1/2 mile or so. This allows for the fact that some people are
more or less sensitive to distances, and that distance is a continuous, and not
a binary, phenomenon. Second, the random utility formulation allows the
researcher to include any attribute of the park; in this case, we consider the
size of the park as an element of accessibility. This suggests that not all
parks are equal, and that a large park such as New York City's Central Park may
provide health and activity benefits over a much larger area than a smaller
community square.

McFadden, D. (1974). The measurement of urban travel demand. Journal of public economics, 3(4), 303-328.

In the absence of a park choice survey, we estimated likely values of the $\beta$ 
coefficients by iteratively searching for the values which produced the highest
model likelihood in *which model did we use to estimate likelihood?*. The search
was constrained by two assertions. First, we required that the coefficient on
size be positive and that on distance negative; all else equal, people will
prefer to use larger parks that are nearer to their residence. Second, we visually
inspected the resulting accessibility scores to ensure that the scores produced a 
reasonably varied pattern of access throughout the metropolitan region.

$$ A_i = log\sum_{j \in J}\exp(distance_{ij} \beta_d + size_j \beta_j + tweets_j \beta_t)$$
We wrote a function to calculate this value for an arbitrary distance matrix, size
and tweets vector, and coefficients.

```{r logsums_function}

#' Calculate destination choice logsums from a distance matrix
#' 
#' @param d An $n\times p$ matrix with the distance from all tracts to
#'   all parks
#' @param sizes A p-length vector of park sizes
#' @param tweets A p-length vector of tweets at parks
#' @param betas A vector containing the size, distance, and tweet coefficients. 
#'   If we submit two variables the tweets are ignored.
#'   
#' @return An n-length vector containing the weighted log-sum based
#'   accessibility between a tract and all parks.
#' @details If we have n tracts and p parks, distances needs to be a 
#' 
calculate_park_logsums <- function(d, sizes, tweets = NULL,
                                   betas = c(-15,.00001, .001)){

  # A is n x p
  a <- betas[1] * d

  # B is p x 1
  b <- betas[2] * sizes

  if(!is.null(tweets)) 
    b <- b + betas[3] * tweets

  # calculate observed utility by adding the weighted park-level attributes 
  # to the columns of the matrix
  # V is n x p, with b added by-column to each element in a
  V <- sweep(a, 2, b, `+`)

  # log-sum of exponentiated utility, Output is n-length vector
  log(rowSums(exp(V)))

}
```

First, we calculate the distance between the tracts and the parks.
```{r distancematrix}
# calculate centroids of tracts and parks
tract_centroids <- tracts %>% st_as_sf() %>% st_centroid() 
park_centroids <- parks %>% st_centroid() %>% st_transform(st_crs(tract_centroids))

# distances from all tracts to all zones
distances <- st_distance(tract_centroids, park_centroids, byid = TRUE) %>%
  units::set_units(miles) %>%
  units::drop_units()
```

Now we want to generate an experiment where we supply different $\beta$ values
to the logsum calculation in order to maximize the log-likelihood function of the
regression model. This will help us identify the proper coefficients to use for
the logsum construction.

## Model without Tweets

```{r notweets_logsum}
notweets_logsum <- function(betas, distances, parks, tracts, model) {

  estimation_data <- tracts@data %>% 
    mutate(logsum = calculate_park_logsums(
      d = distances,  parks$log_size,
      betas = betas))

  sdm <- lagsarlm(update(model, . ~ . + logsum), data = estimation_data, 
                  listw = W, zero.policy = TRUE, type = "mixed")
  return(-as.numeric(logLik(sdm)))
}

access_sdm_val <- optim(
  fn = notweets_logsum,  
  par = c(-1, 0), # distance, size 
  lower = c(-100, -100), # size coefficient must be positive 
  upper = c(0, 100), # distance coefficient must be negative 
  hessian=TRUE, distances = log(distances), parks = parks, tracts = tracts, 
  model = obese_base_lm, method = "L-BFGS-B", control = list(trace = 6))
```

```{r tweets_logsum}
tweets_logsum <- function(betas, distances, parks, tracts, model) {

  estimation_data <- tracts@data %>% 
    mutate(logsum = calculate_park_logsums(
      distances,  parks$log_size,  parks$log_tweets,
      betas = betas))

  sdm <- lagsarlm(update(model, . ~ . + logsum), data = estimation_data, 
                  listw = W, zero.policy = TRUE, type = "mixed")

  return(-as.numeric(logLik(sdm)))
}

tweets_sdm_val <- optim(
  fn = notweets_logsum,  
  par = c(-1, 1, 0),
  lower = c(-100, 0, -100), # size coefficient must be positive
  upper = c(-1e-9, 100, 100), # distance coefficient must be negative
  hessian=TRUE, distances = distances, parks = parks, tracts = tracts, 
  model = obese_base_lm, method = "L-BFGS-B", control = list(trace = 6))

```

```{r obesity_access_models}
tracts@data <- tracts@data %>%
  mutate(
    access_ls = calculate_park_logsums(distances, parks$log_size, 
                                       betas = notweets_lm_val$par),
    tweets_ls = calculate_park_logsums(distances, parks$log_size, parks$log_tweets, 
                                       betas = tweets_lm_val$par),
    # demean
    access_ls = access_ls - mean(access_ls),
    tweets_ls = tweets_ls - mean(tweets_ls)
  )

access_sdm <- lagsarlm(update(obese_base_sdm, .~ . + access_ls),   
                       data = tracts@data,  listw = W, zero.policy = TRUE, 
                       type = "mixed")
tweets_sdm <- lagsarlm(update(obese_base_sdm, .~ . + tweets_ls),   
                       data = tracts@data,  listw = W, zero.policy = TRUE, 
                       type = "mixed")
```

```{r modelchart}
obesity_models <- list(Base = obese_base_sdm, `Access` = access_sdm, 
                       `Access + Tweets` = tweets_sdm)
screenreg(obesity_models, digits = 4)
```

```{r impacts_extractor}
#' Get the MC simulated impact coefficients and significance from a model
#' @param sdm An autoregressive lag (SAR or SDM) model object `lagsarlm`
#' @return A data frame with the effect (direct, indirect, total), variable, 
#'   simulated mean impact, and significance p-value
impacts_extractor <- function(sdm) {

  impacts_summary <- summary(
    impacts(sdm, tr=trMC,  R = 1000, useHESS = !is.logical(sdm$fdHess)), 
    zstats = TRUE)

  coef <- list(
    Direct = impacts_summary$direct_sum,
    Indirect = impacts_summary$indirect_sum,
    Total = impacts_summary$total_sum
  ) %>%
  lapply(function(s) {
    data_frame(
      var = names(s$statistics[,1]),
      impact = s$statistics[,1]
    )
  }) %>%
    bind_rows(.id = "effect")

  pval <- as_data_frame(impacts_summary$pzmat) %>%
    mutate(
     var = rownames(impacts_summary$pzmat)
    ) %>%
    gather("effect", "p-val", -var)

  left_join(coef, pval, by = c("effect", "var"))
}
```

```{r obesity_impacts}
obesity_impacts <- obesity_models %>%
  lapply(impacts_extractor) %>%
  bind_rows(.id = "model") %>%
  transmute(
    model, var, effect, 
    output = paste(round(impact, 5), 
                   gtools::stars.pval(`p-val`))
  ) %>%
  spread(model, output) 
```

# Physical Activity

We cannot necessarily assume that the logsum parameter will be constructed the
same way for physical activity as for obesity. (maybe?) But that's the assumption
we'll make for now.

```{r physact_estimation}
pa_base_sdm <- lagsarlm(update(base_formula, PHYSACT ~ .),
         data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
pa_access_sdm <- lagsarlm(update(base_formula, PHYSACT ~ . + notweets_ls),  
          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
pa_tweets_sdm <- lagsarlm(update(base_formula, PHYSACT ~ . + tweets_ls),  
          data = tracts@data,  listw = W, zero.policy = TRUE, type = "mixed")
```

```{r physact_models}
physact_models <- list(Base = pa_base_sdm, Access = pa_notweets_sdm, 
                       `Access + Tweets` = pa_tweets_sdm)
screenreg(physact_models, digits = 4)
```

```{r physact_impacts}
physact_impacts <- physact_models %>%
  lapply(impacts_extractor) %>%
  bind_rows(.id = "model") %>%
  transmute(
    model, var, effect, 
    output = paste(round(impact, 5), 
                   gtools::stars.pval(`p-val`))
  ) %>%
  spread(model, output) 
```

# Results
```{r impacts_table}
list("Obesity" = obesity_impacts, "Physical activity" = physact_impacts) %>%
  bind_rows(.id = "Outcome") %>%
  filter(var %in% c("access_ls", "tweets_ls")) %>%
  dplyr::select(-var, -Base)  %>%
  gather("Variable", "Impact", -Outcome, -effect) %>%
  filter(!is.na(Impact))  %>%
  spread(effect, Impact)

```

```{r logsum_scales}
data_frame(
  Parameter = c("Distance", "Size", "Tweets"),
  Access = c(notweets_lm_val$par, NA),
  `Access + Tweets` = tweets_lm_val$par
)

```