### 2019-12-16 analysis/sdm.Rmd
make the code visiable

---
output: github_document
---

<!-- # Obesity model estimation -->

Create a spatial weights matrix. It's showing that we have three regions with
no links, so we need to ensure that these are estimated in the model.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(geojsonio)
library(tidyverse)
library(maptools)
library(rgdal)
library(rgeos)
library(spdep)
library(texreg)
library(leaflet)
library(sf)
library(VGAM)
```

```{r W}
# tracts %>% class
# neighbors <- poly2nb(tracts, queen = TRUE)
# neighbors %>% write_rds("../output/neighbors.rds")
neighbors <- read_rds("../output/neighbors.rds")
neighbors %>% class
W <- nb2listw(neighbours = neighbors, zero.policy = TRUE)
print(W, zero.policy = TRUE)
trMC <- trW(as(W, "CsparseMatrix"), type="MC") # trace used in montecarlo impacts
```

```{r obese_lm, echo=FALSE, message=FALSE, include=FALSE}
base_formula <- formula( 
  ~ log(Pop_Density) + 
    FulltimeWork + CollegeDeg + Single_Percent + 
    Pct0to17 + Pct18to29 + Pct65plus + # need to leave out a category for collinearity  
    PctBlack + PctAsian + PctOther + PctHispanic)
# leave out y
parks_formula_1 <- update(base_formula, . ~ . + park_ls1 )
parks_formula_2 <- update(base_formula, . ~ . + park_ls2 )
parks_formula_4 <- update(base_formula, . ~ . + park_ls4 )
parks_formula_5 <- update(base_formula, . ~ . + park_ls5 )
```

First, let's estimate the base model as an LM, SDM, and SEM model

```{r}
# tracts@data %>% write_rds("../output/tracts_data.rds")
data <- read_rds("../output/tracts_data.rds")
data %>% head()
```

```{r basemodels}
# Time consuming.
obese_base_lm <- lm(update(base_formula, OBESITY ~ .), data = data)
obese_base_sar <- lagsarlm(update(base_formula, OBESITY ~ .), 
                           data = data, listw = W, zero.policy = TRUE)
obese_base_sem <- errorsarlm(update(base_formula, OBESITY ~ .), 
                             data = data, listw = W, zero.policy = TRUE)
obese_base_sdm <- lagsarlm(update(base_formula, OBESITY ~ .), 
                           data = data, listw = W, zero.policy = TRUE, 
                           type = "mixed")
```

A likelihood ratio test reveals that the SEM is not preferred, so we use the SDM
only going forward.

```{r lrtest}
lmtest::lrtest(obese_base_sem, obese_base_sdm)
```

```{r obesity, echo=FALSE, message=FALSE, warning=FALSE}
ls_names <- c("Short, no tweets",  "Short, tweets",  
              "Long, no tweets",  "Long, tweets")
# Time consuming.
obese_ls1_sdm <- lagsarlm(update(parks_formula_1, OBESITY ~ .), 
                          data = data,  listw = W, zero.policy = TRUE, type = "mixed")
obese_ls2_sdm <- lagsarlm(update(parks_formula_2, OBESITY ~ .), 
                          data = data,  listw = W, zero.policy = TRUE, type = "mixed")
obese_ls4_sdm <- lagsarlm(update(parks_formula_4, OBESITY ~ .), 
                          data = data,  listw = W, zero.policy = TRUE, type = "mixed")
obese_ls5_sdm <- lagsarlm(update(parks_formula_5, OBESITY ~ .), 
                          data = data,  listw = W, zero.policy = TRUE, type = "mixed")

obesity_models <- list(obese_base_sdm, obese_ls1_sdm, obese_ls2_sdm,
                       obese_ls4_sdm, obese_ls5_sdm)
names(obesity_models) <- c("Base", ls_names)
```

Estimate spatial models for the `Phys_Act` variable. Model 1 does not include park access. Models 2 & 3 each use a different park access variable with a different combination of parameters:

```{r}
screenreg(obesity_models)
```

Estimate the impacts for `obese_sdm_age`.

```{r}
obesity_models %>% class
obesity_models
```

```{r obese_sdm_age_impacts}
# add2pkg::create_desc()
devtools::load_all()
obesity_models %>%
  lapply(impacts_extractor) %>%
  bind_rows(.id = "model") %>%
  transmute(model, var, effect,
            output = paste(round(impact, 5),
                           gtools::stars.pval(`p-val`))) %>%
  spread(model, output)
```

<!-- # Estimate Physical Activity Models -->

```{r phy, eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
physact_base_sdm <- lagsarlm(update(base_formula, Phys_Act ~ .), 
                             data = data,  listw = W, zero.policy = TRUE, 
                             type = "mixed")
physact_ls1_sdm <- lagsarlm(update(parks_formula_1, Phys_Act ~ .), 
                            data = data,  listw = W, zero.policy = TRUE, 
                            type = "mixed")
physact_ls2_sdm <- lagsarlm(update(parks_formula_2, Phys_Act ~ .), 
                            data = data,  listw = W, zero.policy = TRUE, 
                            type = "mixed")
physact_ls4_sdm <- lagsarlm(update(parks_formula_4, Phys_Act ~ .), 
                            data = data,  listw = W, zero.policy = TRUE, 
                            type = "mixed")
physact_ls5_sdm <- lagsarlm(update(parks_formula_5, Phys_Act ~ .), 
                            data = data,  listw = W, zero.policy = TRUE, 
                            type = "mixed")

physact_models <- list(physact_base_sdm, physact_ls1_sdm, physact_ls2_sdm,
                       physact_ls4_sdm, physact_ls5_sdm)
names(physact_models) <- c("Base", ls_names)
```

```{r physact_impacts, eval=FALSE, include=FALSE}
physact_models %>%
  lapply(impacts_extractor) %>%
  bind_rows(.id = "model") %>%
  transmute(
    model, var, effect, 
    output = paste(round(impact, 5), 
                   gtools::stars.pval(`p-val`))
  ) %>%
  spread(model, output) %>% View
```